---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Keian Freshwater"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```


# Simple web scraping

R can read html using either rvest, xml, or xml2 packages. Here we are going to navigate to the Center for Snow and Avalance Studies  [Website](https://snowstudies.org/archived-data/) and read a table in. This table contains links to data we want to programatically download for three sites. We don't know much about these sites, but they contain incredibly rich snow, temperature, and precip data. 


## Reading an html 

### Extract CSV links from webpage

```{r}
site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

#See if we can extract tables and get the data that way
tables <- webpage %>%
  html_nodes('table') %>%
  magrittr::extract2(3) %>%
  html_table(fill = TRUE)
#That didn't work, so let's try a different approach

#Extract only weblinks and then the URLs!
links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('24hr',.)] %>%
  html_attr('href')
```

## Data Download

### Download data in a 'for' loop

```{r}

#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

#generate a file list for where the data goes
file_names <- paste0('data/',dataset)

for(i in 1:3){
  download.file(links[i],destfile=file_names[i])
}

downloaded <- file.exists(file_names)

evaluate <- !all(downloaded)

```


### Download data in a map

```{r}

#Map version of the same for loop (downloading 3 files)
if(evaluate == T){
  map2(links[1:3],file_names[1:3],download.file)
}else{print('data already downloaded')}

```

## Data read-in 

### Read in just the snow data as a loop

```{r}
#Pattern matching to only keep certain files
snow_files <- file_names %>%
  .[!grepl('SG_24',.)] %>%
  .[!grepl('PTSP',.)]

empty_data <- list()

snow_data <- for(i in 1:length(snow_files)){
 empty_data[[i]] <- read_csv(snow_files[i]) %>%
 select(Year,DOY,Sno_Height_M)
 }

snow_data_full <- do.call('rbind',empty_data)

summary(snow_data_full)
```


### Read in the data as a map function

```{r, warning=F}

our_snow_reader <- function(file){
  name = str_split_fixed(file,'/',2)[,2] %>%
    gsub('_24hr.csv','',.)
  df <- read_csv(file) %>%
    select(Year,DOY,Sno_Height_M) %>%
    mutate(site = name)
}

snow_data_full <- map_dfr(snow_files,our_snow_reader)

summary(snow_data_full)
```


### Plot snow data

```{r,warning=F}
snow_yearly <- snow_data_full %>%
  group_by(Year,site) %>%
  summarize(mean_height = mean(Sno_Height_M,na.rm=T))

ggplot(snow_yearly,aes(x=Year,y=mean_height,color=site)) + 
  geom_point() +
  ggthemes::theme_few() + 
  ggthemes::scale_color_few()
```

(DELETE ALL ABV WHEN FINISHED)
# Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r}
#Extracting only the weblinks and then the URLs!

site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

links_forcing <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')

```


2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r}
#Grab only the name of the file by splitting out on forward slashes
splits_forcing <- str_split_fixed(links_forcing,'/',8)

#Keep only the 8th column
dataset_forcing <- splits_forcing[,8] 

#generate a file list for where the data goes
file_names_forcing <- paste0('data/',dataset_forcing)

for(i in 1:2){
  download.file(links_forcing[i],destfile=file_names_forcing[i])
}

downloaded_forcing <- file.exists(file_names_forcing)

evaluate_forcing <- !all(downloaded_forcing)
```

3. Write a custom function to read in the data and append a site column to the data. 

```{r,warning=F,message=F}

# headers from metadata pdf file

library(pdftools)
headers_forcing <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")

#writing the function
forcing_reader <- function(file_names_forcing){
  name = str_split_fixed(file_names_forcing,'/',2)[,2] %>%
    gsub('_Forcing_Data.txt','',.)
  reading = read.delim(file_names_forcing, header = FALSE, sep ="",col.names = headers_forcing,skip = 4) %>%
  mutate(site=name)
}

  
```


4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r}
forcing_data_full <- map_dfr(file_names_forcing,forcing_reader)

summary(forcing_data_full)
```

5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

```{r,warning=F}
forcing_yearly <- forcing_data_full %>%
  group_by(year,site) %>%
  summarize(yearly_air_temp_k = mean(air.temp..K.,na.rm=T))

ggplot(forcing_yearly,aes(x=year,y=yearly_air_temp_k,color=site)) + 
  geom_line() +
  scale_x_continuous(breaks = c(2003,2004,2005,2006,2007,2008,2009,2010,2011))+
  ggthemes::theme_few() + 
  ggthemes::scale_color_few() +
  ylab('Air Temperature (K)')
```

For both the SASP and SBSP sites, here is a sharp increase in air temperature between the years 2003 and 2005; this is at odds with the stable trendline from 2005 onwards. The following graph filters the years before 2005:
```{r,warning=F}
forcing_yearly_filter <- forcing_yearly %>%
  filter(year >= 2005)

ggplot(forcing_yearly_filter,aes(x=year,y=yearly_air_temp_k,color=site)) + 
  geom_line() +
  scale_x_continuous(breaks = c(2005,2006,2007,2008,2009,2010,2011)) +
  ggthemes::theme_few() + 
  ggthemes::scale_color_few() +
  ylab('Air Temperature (K)')
```

6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html
```{r,warning=F,message=F}

forcing_monthly_air_temperature <- forcing_data_full %>%
  group_by(year,month,site) %>%
  summarize(monthly_air_temp_k = mean(air.temp..K.,na.rm=T))


temp_plot <- function(filter_year_function,year_air){
    filter_year_function <- forcing_monthly_air_temperature %>%
      group_by(year,month,site)  %>%
      filter(year==year_air)
    
    str1 = 'Month ('
    str2 = year_air
    str3 = "Year)"
    
    x= ggplot(filter_year_function, aes(x=month,y=monthly_air_temp_k,color=site)) +
    geom_line() +
    scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12), label = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec"))+
    ggthemes::theme_few() +
    ggthemes::scale_color_few() +
    ylab('Air Temperature (K)') +
    xlab(paste(str1,str2,str3))
    
    print(x)
 }

for (year_air in 2005:2010) {
  temp_plot(filter_year_function,year_air)
}


```

From 2005 to 2010, at no point in the year is the Senator Beck Study Plot warmer than the Snow Angel Study Plot.

Bonus: Make a plot of average daily precipitation by day of year (averaged across all available years). 
```{r}

```


Bonus #2: Use a function and for loop to create yearly plots of precipitation by day of year.
```{r}

```
