---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Keian Freshwater"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```


# Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r}
#Extracting only the weblinks and then the URLs!

site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

links_forcing <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')

```


2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r}
#Grab only the name of the file by splitting out on forward slashes
splits_forcing <- str_split_fixed(links_forcing,'/',8)

#Keep only the 8th column
dataset_forcing <- splits_forcing[,8] 

#generate a file list for where the data goes
file_names_forcing <- paste0('data/',dataset_forcing)

for(i in 1:2){
  download.file(links_forcing[i],destfile=file_names_forcing[i])
}

downloaded_forcing <- file.exists(file_names_forcing)

evaluate_forcing <- !all(downloaded_forcing)
```

3. Write a custom function to read in the data and append a site column to the data. 

```{r,warning=F,message=F}

# headers from metadata pdf file

library(pdftools)
headers_forcing <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")

#writing the function
forcing_reader <- function(file_names_forcing){
  name = str_split_fixed(file_names_forcing,'/',2)[,2] %>%
    gsub('_Forcing_Data.txt','',.)
  reading = read.delim(file_names_forcing, header = FALSE, sep ="",col.names = headers_forcing,skip = 4) %>%
  mutate(site=name)
}

  
```


4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r}
forcing_data_full <- map_dfr(file_names_forcing,forcing_reader)

summary(forcing_data_full)
```

5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

```{r,warning=F,message=F}
forcing_yearly <- forcing_data_full %>%
  group_by(year,site) %>%
  summarize(yearly_air_temp_k = mean(air.temp..K.,na.rm=T))

ggplot(forcing_yearly,aes(x=year,y=yearly_air_temp_k,color=site)) + 
  geom_line() +
  scale_x_continuous(breaks = c(2003,2004,2005,2006,2007,2008,2009,2010,2011))+
  ggthemes::theme_few() + 
  ggthemes::scale_color_few() +
  ylab('Air Temperature (K)')
```

For both the SASP and SBSP sites, here is a sharp increase in air temperature between the years 2003 and 2005; this is at odds with the stable trendline from 2005 onwards. The following graph filters the years before 2005:
```{r,warning=F,message=F}
forcing_yearly_filter <- forcing_yearly %>%
  filter(year >= 2005)

ggplot(forcing_yearly_filter,aes(x=year,y=yearly_air_temp_k,color=site)) + 
  geom_line() +
  scale_x_continuous(breaks = c(2005,2006,2007,2008,2009,2010,2011)) +
  ggthemes::theme_few() + 
  ggthemes::scale_color_few() +
  ylab('Air Temperature (K)')
```

6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html
```{r,warning=F,message=F}

forcing_monthly_air_temperature <- forcing_data_full %>%
  group_by(year,month,site) %>%
  summarize(monthly_air_temp_k = mean(air.temp..K.,na.rm=T))


temp_plot <- function(filter_year_function,year_air){
    filter_year_function <- forcing_monthly_air_temperature %>%
      group_by(year,month,site)  %>%
      filter(year==year_air)
    
    str1 = 'Month ('
    str2 = year_air
    str3 = "Year)"
    
    x= ggplot(filter_year_function, aes(x=month,y=monthly_air_temp_k,color=site)) +
    geom_line() +
    scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12), label = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec"))+
    ggthemes::theme_few() +
    ggthemes::scale_color_few() +
    ylab('Air Temperature (K)') +
    xlab(paste(str1,str2,str3))
    
    print(x)
 }

for (year_air in 2005:2010) {
  temp_plot(filter_year_function,year_air)
}


```

From 2005 to 2010, at no point in the year is the Senator Beck Study Plot warmer than the Snow Angel Study Plot.

**Bonus:** Make a plot of average daily precipitation by day of year (averaged across all available years). 
```{r,message=F,warning=F}
forcing_precip_filter <- forcing_data_full %>%
  group_by(year,month,day,site) %>%
  filter(site=='SBB_SASP') %>%
  unite("date", year, month, day, sep = "-", remove = FALSE) %>%
  mutate(date = ymd(date)) %>% 
  mutate(doy = yday(date)) %>%
  summarize(precip= precip..WMO.corrected..kg.m.2.s.1.,
            doy=doy)

forcing_precip_filter_avg <- forcing_precip_filter %>%
  group_by(doy,site) %>%
  summarize(doy_precip = mean(precip))
  
  ggplot(forcing_precip_filter_avg, aes(x=doy,y=doy_precip)) +
    geom_line() +
    ggthemes::theme_few() +
    ggthemes::scale_color_few() +
    ylab('Average Corrected WMO precipitation (kg/m^2*s)') +
    xlab('Day of the Year')
    
```



These graphs show the corrected WMO precipitation in kg/m&2*s for each day of the year, averaged over the data range 2003-2011. Both SBB_SASP and SBB_SBSP report from the same gauge.


**Bonus #2:** Use a function and for loop to create yearly plots of precipitation by day of year.
```{r,message=F,warning=F}
precip_plot <- function(filter_year_function_precip,year_precip){
  
  forcing_precip_function_year<- forcing_data_full %>%
  group_by(year,month,day,site) %>%
  filter(site=='SBB_SASP',
         year==year_precip) %>%
  unite("date", year, month, day, sep = "-", remove = FALSE) %>%
  mutate(date = ymd(date)) %>% 
  mutate(doy = yday(date)) %>%
  summarize(precip= precip..WMO.corrected..kg.m.2.s.1.,
            doy=doy)

  forcing_precip_filter_avg_year <- forcing_precip_function_year %>%
  group_by(doy,site) %>%
  summarize(doy_precip = mean(precip))
    
    str1 = 'Day of the Year ('
    str2 = year_precip
    str3 = "Year)"
    
    z = ggplot(forcing_precip_filter_avg_year, aes(x=doy,y=doy_precip)) +
    geom_line() +
    ggthemes::theme_few() +
    ggthemes::scale_color_few() +
    ylab('Average Corrected WMO precipitation (kg/m^2*s)') +
    xlab(paste(str1,str2,str3))
    
    print(z)
 }

for (year_precip in 2003:2011) {
  precip_plot(filter_year_function_precip,year_precip)
}
```
  

These graphs show the corrected WMO precipitation in kg/m&2*s for each year in the range 2003-2011 Both SBB_SASP and SBB_SBSP are report from the same gauge.